### 2018-01-30 Draft/Draft.Rmd
First commit

---
author:
  - name: Ursula Laa
    affiliation: Monash University
    address: >
      First line
      Second line
    email: \email{ursula.laa@monash.edu}
    url: http://rstudio.com
  - name: Second Author
    affiliation: Affiliation
title:
  formatted: "Scagnostics and Projection Pursuit"
  # If you use tex in the formatted title, also supply version without
  #plain:     "A Capitalized Title: Something about a Package foo"
  # For running headers, if needed
  #short:     "\\pkg{foo}: A Capitalized Title"
abstract: >
  The abstract of the article.
keywords:
  # at least one keyword must be supplied
  formatted: [scagnostics, tour, projection pursuit]
  #plain:     [keywords, not capitalized, Java]
preamble: >
  \usepackage{amsmath}
output: rticles::jss_article
---

```{r initial, echo = FALSE, cache = FALSE, include = FALSE}
library(knitr)
opts_chunk$set(
  warning = FALSE, message = FALSE, echo = FALSE, 
  fig.path = 'figure/', cache.path = 'cache/', fig.align = 'center', 
  fig.show = 'hold', cache = TRUE, external = TRUE, dev = "pdf",
  fig.height = 5, fig.width = 8, out.width = "\\textwidth"
)
read_chunk('src/main.R')
```

```{r load}
```

# Introduction

A grand tour provides dynamic projections from the high dimensions into some low dimensional representation, in particular also as a
two dimensional scatter plot ADDREF.
The idea of projection persuit can be coupled to the tour algorithm, to guide the tour towards more interesting views of the projected data (the so-called guided tour), by maximising a given projection persuit index ADDREF (Cook, Buja, Cabrera, Hurley, 1995).
The guided tour thus allows the user to view interesting projections (local maxima or minima of given index) as well as their neighborhoods, allowing the user to see the projection in the context of the full parameter space.
Typically projection persuit indices are built to detect deviations of the projected data from a normal distribution, and the nonnormalness is considered equivalent to interestingness of a given projection ADDREF (Cook, Buja, Cabrera, 1993).
Moreover, indices for detecting holes, concentration of mass in the center or skewness have also been designed.

Here we aim to extend the definition of interesting projections and available projection persuit indices by studying potential
new index functions.
As a starting point we consdier the scagnostics measures that were proposed for selecting bivariate scatterplots in high-dimensional datasets ADDREF (??).
Rather than considering the scagnostics measures for all possible bivariate combinations, we want to use one of the measures as an index function for a guided tour of 2 dimensional projections.
We first study the feasability of this approach and discuss potential pitfalls.
We then aim to define more robust measures ... ???

# Scagnostics measures

Eight scagnostics measures have been defined, the definitions are based on the convex and alpha hull as well as the minimum spanning tree.
Give definitions here...

Note that in practice the measures are calculated after binning the data to avoid large computing times.

# Analysis setup

To study the feasablility of using scagnostics measures as an index fuction for guided tours, we first consider the variation of each of the measures, as calculated from the projected scatter plot, when smoothly changing the projection via the tour algorithm.
Moreover, we test the dependence on the sample size and on the outlier removal typically performed to make the scagnostics measures more robust.
In addition the computing time is also important, as it might limit the dynamic viewing and require to first record a guided tour path which can
later be viewed in a smooth display.

As we will see the scagnostics measures often exhibit a significant jump despite changing smoothly between projections, a consequence of the discreete nature of the definition, but possibly also due to the binning performed as a first step of calculation.
We therefore explore options of ameliorating this situation using different smoothing methods, and by redefining the measures in a more robust fashion.

# Application to particle physics data

## Data description

For the following illustrations we work with the particle physics dataset (ATLAS Collaboration, 2015).
This data includes a number of predictions for observables (in agreement with current limits) and the so-called finetuning measure for a large set of parameter points in a supersymmetric model.
The observables describe predictions for decay branching ratios and other precision observables that are generally used to constrain new models of particle physics.
We use the log transformed values for the variables XYZ but not the finetuning, and we rescale all variables to a [0,1] interval
to obtain comparable scales.
Contributions to the variable gminus2 can be negative or positive, we therefore add a small shift before the log transformation of this variable.

## Software setup

Use tourr and scagnostics package (ADDREF)

## Scagnositics of grand tour projections

```{r scagnostics-summary, fig.height= 14, fig.width=9, fig.cap = "(ref:scagnostics-summary)"}
```

(ref:scagnostics-summary) Summary of evolution of scagnostics measures over 100 interpolated grand tour projections
We see that three of the considered measures, Clumpy, Sparse and Striated, are not relevant for the considered data and projections,
as the valuse are consistently low (clearly below 0.1).
The remaining measures show large variation depending on the projection, sample size and also often strong dependence on the
outlier removal before calculation of scagnostics measures.
Moreover, we note that despite considering interpolated projections, i.e. a smooth transition between the randomly selected projection planes in the grand tour, the variation of the scagnostics measures does not follow a smooth evolution but exhibits several spikes.
This behaviour is expected to result in difficulites when using scagnostics measures as a projection pursuit index in the context of a dynamic guided tour.
To further understand this behaviour let us study the projected data points and also the binned version for some of the projections exhibiting a spike.
We first show in Figure ADDREF projections t=18/19/20 of the small data set (500 data points).
Not that t=19 has been marked by a vertical line in Figure ADDREF.
The left column shows the projected data points, and the right column the binned data used to calculate the scagnositcs measures.
The scagnostics measures in the three projections are summarised in Table ADDREF.
We see that despite similar distribution of the points in all three projections, the outlying measure is similar for t=18 and t=20 but jumpig down for t=19.
Differences can however be observed in the binned data (right column of Figure ADDREF), likely causing the behaviour of the measures.

(This is blueprint of how to show scatter and binning plots but maybe other projections will be more interesting.)

FIXME need to arrange plots better! Fix common axes for all corresponding plots
FIXME binned plot needs better formatting so we can see relevant features easily

```{r scatter-binning, fig.height=2, fig.cap = "(ref:scatter-binning)"}
```

```{r scatter-table}
```

Another consideration relevant for the purpose of the guided tour is the computing time, as mentioned above it may limit the tour display to previously recorded tours.
We study here the computing time depending on the sample size with and without the outlier removal when calculating the measures.
Figure ADDREF shows this variation for 3 different sample selections.
The time is measured over the 100 projections shown before and is the sum of computing time for projecting the data and calculating the scagnostics measures on the projected data.
We see that overall the outlier removal increases the runtime by about a factor two.
However we have seen above that skipping the step of outlier removal often leads to very different values of some of the scagnostics measures. (What to do with this information? Should not skip this step?)
Interestingly the computing times are longest for intermediate sample sizes of a few hundred data points, and converge to lower run times for larger samples.
This is because the binning of the input data is much faster than the computation of the measures which is increased for smaller samples.

```{r scagnositcs-timing, fig.cap = "(ref:scagnostics-timing)"}
```

## Smooth transitions

We next turn to the question of how we can obtain smooth transitions of the scagnostics measures.
(Below just an example of using ggplot smoothing to display what kind of behaviour we want to find. It follows the general trend without capturing the spikes.)
```{r smoothing, fig.cap = "(ref:smooting)"}
```

Want to explore jittering either the projection or the points, but this does not give the desired smooth behaviour (why?)
Maybe show how it compares to the sketch of smooth polynomial function? Select parameters from previous study without including the detailed plots here..

### Binning in high dimensions?
### Outlier removal before binning?

## Test with guided tour
try out scagnostics measures as guided tour index
FIXME still need to write script for this

## Additional indices to consider
FIXME load Katrins package and try the index functions she implemented