### 2019-06-05 review/response.Rmd
Adding comments from reviewer 2.

COMMENTS FOR THE AUTHOR:

Associate Editor:

The submission has been evaluated by two reviewers. The manuscript is well written and organized. I have some more philosophical issues in addition to those raised by the two reviewers.

The authors need to convince readers (users) this kind of environment is good for daily EDA for the real world of big data/data science.

Example(s) with very large number (at least hundreds) of dimensions should be included in the study.

**Theoretically, it might work for really big data, but there are two components that would be required: (1) fast computational engine, (2) an efficient optimiser. Both of these would require major implementation work. We have explained this now in the discusion section.**

**We have analysed an example that is high-dimensional but think that it would be too distracting for the reader of this paper. It is an analysis of physics modeled fits, using initial ideas developed in this paper. There the dimensionality was around 60, and it worked best to reduce the dimensionality using PCA as a pre-processing step, because this explained most of the variability. The tour was applied on the reduced dimensionality, and the work is published in EPJC (cite).**

Who are the target users of this environment? Do you recommend naive users (users without much EDA experience) to get hand on the proposed system?

How do you encourage users to try your system?

**The target audience consists of two groups. First we address people interested in developping new index functions, who will find the introduced methods useful to diagnose new indexes as presented in the paper. The methods to do so are now easily accessible via the spinebil package. The second group would be people interested in applying the considered indexes to their datasets. This can be done directly using the tourr package, but we also encourage users unfamiliar with the system to try this through the GUI available in the galahr package. Both packages are now referenced in the discussion (or should it be mentioned somewhere else?).**

**The methods are implemented in an R package, which makes them readily available to all. The code for this paper is all available on github, and we have now made this repo public. It depends on three other packages, mbgraphics, binostics, minerva, which are now updated and on CRAN. We have coded the diagnostic techniques into a new package, and will submit this to CRAN.**

Reviewer #1: Review report for "Using tours to visually investigate properties of new projection pursuit indexes with application to problems in physics"

This manuscript presented a simulation study to visually explore the properties of a set of projection pursuit indexes (PPIs) using a grand tour based on the five criterions. The selected PPIs were then applied to two real datasets in physics for revealing potential patterns. In general I found this article was well written and it fully addresses the problems to be solved. I have no major issue. Some specific comments are listed below:

1) Authors conducted the simulation study to evaluate the selected PPIs based on the five criterions and the results were summarized into Table 3. On the other hand, authors may consider another table that summarizes the advantages/disadvantages/characteristics of these mathematical indexes from the previous literatures. This review is helpful for users to comprehend these indexes' properties.

**This comparison was the subject of the thesis referenced in the introduction (Grimm (2016)). We have now added a brief summary of these findings at the end of section 2.2.**

2) According to the aims & scope of the journal, Computational Statistics, this article should keep the content and discussion relevant more focused on the computational aspects.

**maybe move more things to the appendix? not sure how to reply here...**

3) Although the target of this article focuses on the dataset with two-dimensional structure (d=2), it is worth to discuss the possibility (or give some suggestions) of applying the proposed procedure to the situation where d > 2.

**We thank the referee for the suggestion. Indeed much of the methods developped for diagnostics of the index function apply directly to scenarios where d > 2. While the considered index functions may in principle be generalised to higher dimensions as well, this is currently only implemented for the distance correlation function. A detailed assessment of the d > 2 situation will be the subject of future work. (as now stated in the discussion?)**

4) The procedure for assessing the PPIs' properties was listed in Section 3.1.2. In addition, it would be helpful to list the procedure step-by-step (with the detailed R functions/packages) to guide the users how to apply the tour methodology with PPIs to explore the potential patterns of the real-world datasets.

**The functions used for the evaluation of these tests are now packaged in the spinebil R package. The package documentation contains examples to guide users and make it easy to reproduce the results for different test datasets or different index functions.**

5) Why use the 5th/95th percentiles in Table 2 instead of the average/sd?

**The main reason is that it provides a range of typical values of the index, which is useful for scaling the index dynamically in the guided tour. While it is possible to use a mean, and maybe +/- 3sd's for a similar purpose, range is more useful.**

6) There was some confusion regarding the number of PPIs used in the content. There are eight indexes in Table 2; nine indexes in Fig 7; six indexes in Table 3, and seven indices in Section 4.2.2.

**Need to decide treatment of holes/cmass index!**

7) I would suggest moving the appendix A (the description of "holes/cmass" index) to Section 2.2 for readability.

**Mention holes/cmass with reference in Sec. 2.2, but keep rescaling information in the appendix?**

8) Check the spelling of words for clarity and consistency, e.g, splines2D/splines2d; convex/1-convex, MIC/MICe, and so on.

**We have made the spelling consistent where referring to the same index, and provided explanations where the different spelling indicates differences in index calculation.**

Reviewer #2:
Summary of the contribution:
The authors present a simulation study to evaluate various projection pursuit indexes for detecting specific bivariate patterns in the data. Moreover, two data sets from physics are examined using PPI.

Major remarks
The simulation study targets three particular structures the pipe, the sine, and the spiral. Some motivation and justification for selecting these structures should be added.

 **We have added a statement in Section 3.1.1 outlining the motivation for the three structures considered in the simulation study.**

Second, the simulation study uses six-dimensional data of which four dimensions con- tribute nuisance and the other two the intended structure. Implicitly, this is clear, but the description does not make this explicit and uses the general notion of p dimensions instead. Please make this more explicit and also provide some justification that this set-up is adequate for targeting the problem at hand.

**We have added an explicit statement that we fix p=6. As mentioned in the statement p=6 results in two independent planes showing the nuisance distribution and one plane showing the structured distribution. This set-up is adequat for most of our study where we follow interpolated paths either between nuisance only projections or from nuisance views to structured projections.**

Thirdly, the simulations refer to sequences of projections (the horizontal axis in plots in Figures 2, 3, and 4. As I understand these sequences are a consequence of optimising the corresponding projection pursuit index. How can you then actually control the two different scenarios referred to in section 3.2 and 3.3? Or, do you extract a specified sequence of projections? Moreover, how do you make sure that you have the same sequence length for each trial? Please clarify!

**Figures 2 and 3 present the trace for a specified sequence of projections (since in the simulation study the nuisance and structured views are known). This is explained in Section 3.1.2, and we have added a sentence clarifying this in Section 3.2 and made the wording more explicit in Section 3.3. Figure 4 is different as in this case the path is indeed the result of optimising an index. However, we only optimise the TIC index highlighted in red, while the other index values are shown for comparison along the optimisation path.**

**(In my opinion all of this was already very clear before. When discussing this before you suggested making the Figure caption more clear, but it seems they are already very clear and I would not want to make them longer.)**

The optimisation methods are further discussed in section 3.4 and depend on the struc- ture present in the data. What is the value of these investigations for more realistic settings in which the structure is not known.

**(I would list: demonstrate potential as PPI, understand shortcomings, and based on that develop strategies, e.g. the two-step optimisation. Add something like this in Section 3.4?)**

Finally, the relationship between the simulation study and the examples using the astrophysics data remains unclear. How are the physics applications linked to the simulation study and vice versa? Currently, the two aspects seem to me rather independent and could actually be presented in separate manuscripts. Moreover, since some rather important aspects of the simulation study have been put into an appendix. For the COST audience certainly the details of the simulation study are more relevant than the astrophysics application.

**(Point to initial motiviation, what else should we say here?)**

Minor remarks
When introducing the various PPIs some terms and notations are used which are not properly introduced, e.g. alpha hull, convex. Check which ones actually need to be properly defined as they are needed for further steps and for which ones you could just refer to the literature avoiding unnecessary details.

**(Still need to do this)**

In Figure 3, each plot shows a vertical blue line. Please explain what this line represents.

**We have added such a sentence to the figure caption.**

p. 14, l. 26: optimist should be optimise

**We have fixed the typo.**